### 2014-11-09 Introduction.tex
stuff

The goal of this work is to provide a framework for full verification of such
complex hardware systems. Hardware designers already recognize the challenge of
implementing these systems, to such an extent that they have already become some
of the most serious real-world adopters of formal methods.  Hardware
verification is dominated by model-checking, where abstraction techniques are
used to reduce designs to finite state spaces, which can be explored
exhaustively.  There are limits to the construction of sound abstractions, so
verifications of, e.g., cache-coherence protocols have almost always treated
systems with concrete topologies, involving particular finite numbers of caches
and processors. For instance, explicit-state model checking tools like
Murphi~\cite{murphi} or TLC~\cite{tlc} are only able to handle small and
simplified systems, \eg{} a single-level cache hierarchy with three CPUs, two
addresses, and one bit of data. Symbolic model checking techniques have fared
slightly better: McMillan \etal{} have verified a 2-level MSI protocol based on
the Gigamax distributed multiprocessor using SMV~\cite{gigamax}. Optimizations
on these techniques (\eg{} partial order reduction~\cite{part}, symmetry
reduction~\cite{sym1, sym2}, compositional reasoning~\cite{comp, Mccomp, mcc}, \etc{}) scale
the approach further, but they are still unable to handle realistic
hierarchical protocols. We might hope to achieve higher assurance and
understanding of design ideas by verifying \emph{infinite families} of hardware
designs, which resist reduction to finite-state models.  A potential related
benefit is reducing the computational resources required to formally verify a
new design in the family, applying a more general theorem without much new
analysis.

@@

LTSes as hardware descriptions are an established
idea~\cite{HoeArvind:TRSSynthesis1, Hoe:TCAD}, and there are
compilers that convert LTSes into efficient hardware.  Our work is
based on the Bluespec language~\cite{BSV:LangRef, Bluespec:TFRG},
which served as the model for the formalism of this paper.  Bluespec
models hardware components as atomic rules of a transition system over
state elements, and there is a commercial compiler from such code into
register transfer-level circuits (i.e., Verilog code) with competitive
performance.  Our cache-coherent memory system is directly
transliterated from a Bluespec
implementation~\cite{DNA:CoherenceImplementation} used to
implement a multi-processor PowerPC system~\cite{Khan:PowerPc}. The
hardware synthesized from that implementation is rather efficient:
an 8-core system can run 55 million instructions per second on
the BEE FPGA board\cite{}, with a 2-level cache hierarchy.

>>>>

The goal of this work is to provide a framework for full verification of such
complex hardware systems. Hardware designers already recognize the challenge of
implementing these systems, to such an extent that they have already become some
of the most serious real-world adopters of formal methods.  Hardware
verification is dominated by model-checking; for instance, processor
verification ~\cite{burch1994automatic, mcmillan1998verification} and 
more recently, Intel's execution cluster verification~\cite{kaivola2009replacing}.
Many abstraction techniques are
used to reduce designs to finite state spaces, which can be explored
exhaustively.
 There are limits to the construction of sound abstractions, so
verifications of, e.g., cache-coherence protocols have almost always treated
systems with concrete topologies, involving particular finite numbers of caches
and processors. For instance, explicit-state model checking tools like
Murphi~\cite{murphi} or TLC~\cite{tlc} are only able to handle small and
simplified systems, \eg{} a single-level cache hierarchy with three CPUs, two
addresses, and one bit of data. Symbolic model checking techniques have fared
slightly better: McMillan \etal{} have verified a 2-level MSI protocol based on
the Gigamax distributed multiprocessor using SMV~\cite{gigamax}. Optimizations
on these techniques (\eg{} partial order reduction~\cite{part}, symmetry
reduction~\cite{sym1, sym2}, compositional reasoning~\cite{comp, Mccomp, mcc}, \etc{}) scale
the approach further, but they are still unable to handle realistic
hierarchical protocols. We might hope to achieve higher assurance and
understanding of design ideas by verifying \emph{infinite families} of hardware
designs, which resist reduction to finite-state models.  A potential related
benefit is reducing the computational resources required to formally verify a
new design in the family, applying a more general theorem without much new
analysis.

@@

LTSes as hardware descriptions are an established
idea~\cite{HoeArvind:TRSSynthesis1, Hoe:TCAD}, and there are
compilers that convert LTSes into efficient hardware.  Our work is
based on the Bluespec language~\cite{BSV:LangRef, Bluespec:TFRG},
which served as the model for the formalism of this paper.  Bluespec
models hardware components as atomic rules of a transition system over
state elements, and there is a commercial compiler from such code into
register transfer-level circuits (i.e., Verilog code) with competitive
performance.  Our cache-coherent memory system is directly
transliterated from a Bluespec
implementation~\cite{DNA:CoherenceImplementation} used to
implement a multi-processor PowerPC system~\cite{Khan:PowerPc}. The
hardware synthesized from that implementation is rather efficient:
an 8-core system can run 55 million instructions per second on
the BEE FPGA board\cite{}, with a 2-level cache hierarchy. Figure \ref{both} gives an example
of an LTS inference rule transliterated directly from an actual Bluespec rule used
in the PowerPC system.

\begin{figure}[t]
\centering
\begin{boxedminipage}{\columnwidth}
\small
\noindent \textbf{rule} $(c.\parent == p \; \mathit{\&\&} \; c.\s[a] < x \; \mathit{\&\&} \; c.\wt[a]==$None$);$

\hspace{1cm} $\ch[c][p][\req].\enq(a,c.\s[a],x); c.\wt[a] \Leftarrow x;$

\textbf{endrule}
\end{boxedminipage}
\begin{boxedminipage}{\columnwidth}
\small
\inference[]
{\parent(c,p) & \s(c,a) < x & \wt(c,a) = \epsilon}
{\ioxe{$M_c$}
{\dt, \ch, \s,\\ \dst, \wt, \dwt,\\ \inp, \outp}
{\dt, \ch[ (c, p, \req) \coloneqq(a, \s(c, a),x)\\
\;\;\uplus \ch(c,p,\req)], \s, \dst,\\
\wt[(c,a)\coloneqq x], \dwt,\inp, \outp}{}}
\end{boxedminipage}
\caption{Rule in Bluespec and its corresponding LTS transliteration}
\label{both}
\end{figure}

### 2014-11-09 RelatedWork.tex
lots of modifications
stuff

Formal verification has been used to verify multiprocessor
systems, especially with model checking
\cite{burch1994automatic,mcmillan1998verification}.

More recently, Intel's execution cluster (\ie{} the ALU and its
interface) has been formally verified using model checking
\cite{kaivola2009replacing}.

@@

%\murali{Next paragraph is all I know about this stuff, should probably be
%removed}
%\adam{What are you referring to with ``it'' in your comment?  I don't understand what this paragraph refers to.}
%It has been verified using model checking \cite{qadeer2003verifying},
%with a fixed number of processors and memory locations.  Model-checking has also
%been used to verify communication fabrics \cite{chatterjee2012automatic}.

Theorem provers have also been used in verifying cache coherence protocols.
The correctness of the FLASH cache coherence protocol~\cite{flash} for a
single-level hierarchy was proven with PVS~\cite{park} using a mapping, called
``aggregation of transactions,'' from multiple localized transactions to a
single global transaction, and finally using model checking to prove the correctness of
the global transactions model. However, they do not provide a methodology to
prove the aggregation of transactions for other protocols, which is where, we
believe, the difficulty lies; in contrast, our modular proof relies on generic
invariants, making it more adaptable to different protocols.

Cachet~\cite{StoyShenArvind:Proofs}, another single-level cache coherence
protocol, was proven to be implementable in the CRF memory
model~\cite{Shen:CRF}. Shen \etal{} provided a paper-and-pencil proof showing
Cachet does not violate the CRF memory model, while Stoy \etal{} implemented
portions of the paper-and-pencil proof using PVS. However, neither proof
establishes that the protocol is store atomic. Rather, they show that Cachet
obeys the CRF memory model, which is not restricted to store-atomic behavior.

\xxx{TODO: Correct This}

>>>>

%Formal verification has been used to verify multiprocessor
%systems, especially with model checking
%\cite{burch1994automatic,mcmillan1998verification}.

%More recently, Intel's execution cluster (\ie{} the ALU and its
%interface) has been formally verified using model checking
%\cite{kaivola2009replacing}.

@@

Theorem provers have also been used in verifying cache coherence protocols.
The correctness of the FLASH cache coherence protocol~\cite{flash} for a
single-level hierarchy was proven with PVS~\cite{park} using a mapping, called
``aggregation of transactions,'' from multiple localized transactions to a
single global transaction, and finally using model checking to prove the correctness of
the global transactions model. However, they do not provide a methodology to
prove the aggregation of transactions for other protocols, which is where, we
believe, the difficulty lies; in contrast, our modular proof relies on generic
invariants, making it more adaptable to different protocols.
Cachet~\cite{StoyShenArvind:Proofs}, another single-level cache coherence
protocol, was verified to be obey the CRF memory
model~\cite{Shen:CRF}, and parts of the proof were implemented using PVS.

% Shen \etal{} provided a paper-and-pencil proof showing
%Cachet does not violate the CRF memory model, while Stoy \etal{} implemented
%portions of the paper-and-pencil proof using PVS. However, neither proof
%establishes that the protocol is store atomic. Rather, they show that Cachet
%obeys the CRF memory model, which is not restricted to store-atomic behavior.
